When easy is not preferred: A discounting paradigm to assess load-independent task preference

### 2023-09-01 06_Paper/COG-ED/Stage 2/Manuscript.Rmd
included acknowledgement and fixed 3 digit format

---
title             : "When easy is not preferred: A discounting paradigm to assess load-independent task preference"
shorttitle        : "The CAD paradigm to assess task preference"
author: 
  - name          : "Josephine Zerna"
    orcid         : "0000-0003-2892-884X"
    affiliation   : "1"
    corresponding : yes
    address       : "Zellescher Weg 17, 01069 Dresden, Germany"
    email         : "josephine.zerna@tu-dresden.de"
    equal_contrib : yes
    role:         # Contributorship roles (e.g., CRediT, credit.niso.org)
      - Conceptualization
      - Data curation
      - Methodology
      - Funding acquisition
      - Formal analysis
      - Investigation
      - Project administration
      - Software
      - Visualization
      - Writing - original draft
      - Writing - review & editing
  - name          : "Christoph Scheffel"
    orcid         : "0000-0001-5963-9229"
    affiliation   : "1"
    equal_contrib : yes
    role:
      - Conceptualization
      - Methodology
      - Funding acquisition
      - Investigation
      - Project administration
      - Software
      - Writing - review & editing
  - name          : "Corinna Kührt"
    orcid         : "0000-0002-6418-6479"
    affiliation   : "1"
    role:
      - Formal analysis
      - Writing - review & editing
  - name          : "Georgia Clay"
    orcid         : "0000-0001-5641-6804"
    affiliation   : "2"
    role:
      - Formal analysis
  - name          : "Alexander Strobel"
    orcid         : "0000-0002-9426-5397"
    affiliation   : "1"
    role: 
      - Conceptualization
      - Resources
      - Supervision
      - Funding acquistion
      - Writing - review & editing
affiliation:
  - id            : "1"
    institution   : "Faculty of Psychology, Dresden University of Technology, 01062 Dresden, Germany"
  - id            : "2"
    institution   : "Faculty of Psychology, University of Vienna, Vienna 1010, Austria"
authornote        :
abstract          : |
  When individuals set goals, they consider the subjective value (SV) of the anticipated reward and the required effort, a trade-off that is of great interest to psychological research.
  One approach to quantify the SVs of levels of difficulty of a cognitive task is the Cognitive Effort Discounting Paradigm by Westbrook and colleagues (2013).
  However, it fails to acknowledge the highly individual nature of effort, as it assumes a unidirectional, inverse relationship between task load and SVs.
  Therefore, it cannot map differences in effort perception that arise from traits like Need for Cognition, since individuals who enjoy effortful cognitive activities likely do not prefer the easiest level.
  We replicated the analysis of Westbrook and colleagues with an adapted version, the Cognitive and Affective Discounting (CAD) Paradigm.
  It quantifies SVs without assuming that the easiest level is preferred, thereby enabling the assessment of SVs for tasks without objective order of task load.
  Results show that many of the 116 participants preferred a more or the most difficult level.
  Variance in SVs was best explained by a declining logistic contrast of the $n$-back levels and by the accuracy of responses, while reaction time as a predictor was highly volatile depending on the preprocessing pipeline.
  Participants with higher Need for Cognition scores perceived higher $n$-back levels as less effortful and found them less aversive.
  Effects of Need for Cognition on SVs in lower levels did not reach significance, as group differences only emerged in higher levels.
  The CAD Paradigm appears to be well suited for assessing and analysing task preferences independent of the supposed objective task difficulty.

@@

There were three necessary adjustments of the methods.
First, we failed to update the necessary sample size after the analyses changed with the first review round.
Instead of the 72 subjects stated above, the largest minimum sample size was actually 53 subjects (see hypothesis 1b in the Design Table in the Supplementary Material).
Secondly, we changed to which hypothesis we applied the specification curve analysis (SCA).
In the initial Stage 1 submission, we had applied it to the MLM of hypothesis 2b, which at this point included NFC as a predictor.
Following the advice of the reviewers, we removed NFC from the MLM, and analyzed NFC in an rmANOVA (hypothesis 3a) instead.
Since NFC was of great interest to us, we decided to apply the SCA to hypothesis 3a rather than 2b to provide a measure of robustness.
However, hypothesis 3a does not contain any RT data, so the SCA is only useful for the MLM in hypothesis 2b.
Therefore, we applied it to the MLM.
The final adjustment was made during the Stage 2 revision.
A fellow researcher made us aware that by using the z-transformed hit and false alarm rates for the computation of $d'$, the mean of $d'$ would be approximately 0 for each $n$-back level by design.
Consequently, $d'$ could not show changes across $n$-back levels in the manipulation check and would likely yield different results in the MLM.
Therefore, we computed $d'$ with unstandardized hit and false alarm rates.
We would like to thank G.C. for pointing us to this fallacy and included her as a co-author.

@@

The declining logistic $n$-back levels and $d'$ remained significant predictors of SVs throughout all 63 preprocessing pipelines in the specification curve analysis.
The effect estimates of the $n$-back level varied by about `r round(diff(range(sca_results$beta_n)), digits = 3)`, those of $d'$ by about `r round(diff(range(sca_results$beta_d)), digits = 3)`.
In contrast to this stood the variability of the median RT effect estimates (around `r round(diff(range(sca_results$beta_r)), digits = 3)`), which did not reach significance in any pipeline.
Interestingly, the curve of median RT betas in Figure\ \@ref(fig:sca-plot)a roughly mirrored the rectangular pipeline indicators in the transformation rows of Figure\ \@ref(fig:sca-plot)b, so the transformation choice influenced the median RT much more than the dimension or the exclusion choice did.
As Fernandez et al.[-@Fernandez2020] found, applying more than one preprocessing step to the reaction time data of a Stroop task increased the risk of false positives beyond $\alpha=.05$, and transformation choices inflated this risk more than outlier exclusion or aggregation choices did.
Our data seems to corroborate this finding for $n$-back tasks as well.
Surprisingly, the $d'$ betas appeared almost unaffected by the preprocessing pipeline, even though $d'$ was computed after the outlier exclusion.
This indicates that researchers who are interested in the correctness rather than the speed of responses can choose a simple preprocessing pipeline without risking false positives through elaborate transformations.

@@

This research was partly funded by the German Research Foundation (DFG) as part of the Collaborative Research Center (CRC) 940, and partly funded by centralized funds of the Faculty of Psychology at Technische Universität Dresden.
The funders have/had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.
The authors would like to thank Juliana Krause and Maja Hentschel for their help with data collection.

@@

JZ and CS contributed equally to this work.
JZ, CS, and AS conceptualized the study and acquired funding.
JZ and CS developed the methodology, investigated, administered the project, and wrote the software.
JZ, CS, GC, and CK did the formal analysis.
JZ visualized the results.
JZ and CK prepared the original draft.
All authors reviewed, edited, and approved the final version of the manuscript.

>>>>

---
title             : "When easy is not preferred: A discounting paradigm to assess load-independent task preference"
shorttitle        : "The CAD paradigm to assess task preference"
author: 
  - name          : "Josephine Zerna"
    orcid         : "0000-0003-2892-884X"
    affiliation   : "1"
    corresponding : yes
    address       : "Zellescher Weg 17, 01069 Dresden, Germany"
    email         : "josephine.zerna@tu-dresden.de"
    equal_contrib : yes
    role:         # Contributorship roles (e.g., CRediT, credit.niso.org)
      - Conceptualization
      - Data curation
      - Methodology
      - Funding acquisition
      - Formal analysis
      - Investigation
      - Project administration
      - Software
      - Visualization
      - Writing - original draft
      - Writing - review & editing
  - name          : "Christoph Scheffel"
    orcid         : "0000-0001-5963-9229"
    affiliation   : "1"
    equal_contrib : yes
    role:
      - Conceptualization
      - Methodology
      - Funding acquisition
      - Investigation
      - Project administration
      - Software
      - Writing - review & editing
  - name          : "Corinna Kührt"
    orcid         : "0000-0002-6418-6479"
    affiliation   : "1"
    role:
      - Formal analysis
      - Writing - review & editing
  - name          : "Alexander Strobel"
    orcid         : "0000-0002-9426-5397"
    affiliation   : "1"
    role: 
      - Conceptualization
      - Resources
      - Supervision
      - Funding acquistion
      - Writing - review & editing
affiliation:
  - id            : "1"
    institution   : "Faculty of Psychology, Dresden University of Technology, 01062 Dresden, Germany"
authornote        :
abstract          : |
  When individuals set goals, they consider the subjective value (SV) of the anticipated reward and the required effort, a trade-off that is of great interest to psychological research.
  One approach to quantify the SVs of levels of difficulty of a cognitive task is the Cognitive Effort Discounting Paradigm by Westbrook and colleagues (2013).
  However, it fails to acknowledge the highly individual nature of effort, as it assumes a unidirectional, inverse relationship between task load and SVs.
  Therefore, it cannot map differences in effort perception that arise from traits like Need for Cognition, since individuals who enjoy effortful cognitive activities likely do not prefer the easiest level.
  We replicated the analysis of Westbrook and colleagues with an adapted version, the Cognitive and Affective Discounting (CAD) Paradigm.
  It quantifies SVs without assuming that the easiest level is preferred, thereby enabling the assessment of SVs for tasks without objective order of task load.
  Results show that many of the 116 participants preferred a more or the most difficult level.
  Variance in SVs was best explained by a declining logistic contrast of the $n$-back levels and by the accuracy of responses, while reaction time as a predictor was highly volatile depending on the preprocessing pipeline.
  Participants with higher Need for Cognition scores perceived higher $n$-back levels as less effortful and found them less aversive.
  Effects of Need for Cognition on SVs in lower levels did not reach significance, as group differences only emerged in higher levels.
  The CAD Paradigm appears to be well suited for assessing and analysing task preferences independent of the supposed objective task difficulty.

@@

There were three necessary adjustments of the methods.
First, we failed to update the necessary sample size after the analyses changed with the first review round.
Instead of the 72 subjects stated above, the largest minimum sample size was actually 53 subjects (see hypothesis 1b in the Design Table in the Supplementary Material).
Secondly, we changed to which hypothesis we applied the specification curve analysis (SCA).
In the initial Stage 1 submission, we had applied it to the MLM of hypothesis 2b, which at this point included NFC as a predictor.
Following the advice of the reviewers, we removed NFC from the MLM, and analyzed NFC in an rmANOVA (hypothesis 3a) instead.
Since NFC was of great interest to us, we decided to apply the SCA to hypothesis 3a rather than 2b to provide a measure of robustness.
However, hypothesis 3a does not contain any RT data, so the SCA is only useful for the MLM in hypothesis 2b.
Therefore, we applied it to the MLM.
The final adjustment was made during the Stage 2 revision.
A fellow researcher made us aware that by using the z-transformed hit and false alarm rates for the computation of $d'$, the mean of $d'$ would be approximately 0 for each $n$-back level by design.
Consequently, $d'$ could not show changes across $n$-back levels in the manipulation check and would likely yield different results in the MLM.
Therefore, we computed $d'$ with unstandardized hit and false alarm rates.
We would like to thank Georgia Clay for pointing us to this fallacy.

@@

The declining logistic $n$-back levels and $d'$ remained significant predictors of SVs throughout all 63 preprocessing pipelines in the specification curve analysis.
The effect estimates of the $n$-back level varied by about `r format(round(diff(range(sca_results$beta_n)), digits = 3), digits = 3)`, those of $d'$ by about `r format(round(diff(range(sca_results$beta_d)), digits = 3), digits = 3)`.
In contrast to this stood the variability of the median RT effect estimates (around `r round(diff(range(sca_results$beta_r)), digits = 3)`), which did not reach significance in any pipeline.
Interestingly, the curve of median RT betas in Figure\ \@ref(fig:sca-plot)a roughly mirrored the rectangular pipeline indicators in the transformation rows of Figure\ \@ref(fig:sca-plot)b, so the transformation choice influenced the median RT much more than the dimension or the exclusion choice did.
As Fernandez et al.[-@Fernandez2020] found, applying more than one preprocessing step to the reaction time data of a Stroop task increased the risk of false positives beyond $\alpha=.05$, and transformation choices inflated this risk more than outlier exclusion or aggregation choices did.
Our data seems to corroborate this finding for $n$-back tasks as well.
Surprisingly, the $d'$ betas appeared almost unaffected by the preprocessing pipeline, even though $d'$ was computed after the outlier exclusion.
This indicates that researchers who are interested in the correctness rather than the speed of responses can choose a simple preprocessing pipeline without risking false positives through elaborate transformations.

@@

This research was partly funded by the German Research Foundation (DFG) as part of the Collaborative Research Center (CRC) 940, and partly funded by centralized funds of the Faculty of Psychology at Technische Universität Dresden.
The funders have/had no role in study design, data collection and analysis, decision to publish or preparation of the manuscript.
The authors would like to thank Juliana Krause and Maja Hentschel for their help with data collection, and Georgia Clay for noticing and calling our attention to the fallacy of z-transforming the hit and false alarm rates.

@@

JZ and CS contributed equally to this work.
JZ, CS, and AS conceptualized the study and acquired funding.
JZ and CS developed the methodology, investigated, administered the project, and wrote the software.
JZ, CS, and CK did the formal analysis.
JZ visualized the results.
JZ and CK prepared the original draft.
All authors reviewed, edited, and approved the final version of the manuscript.
